---
title: "Use of Force Final Project Report"
author: "Kaylhan Garcia, Adam Billen, and Phoebe Lin"
date: "2023-10-17"
output:
  pdf_document: default
  word_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
library(conflicted)
library(dplyr)
dplyr::filter
conflicts_prefer(dplyr::filter)
```

# Use of Force Final Project Report

**Introduction:**

Police have existed in America in some form for nearly its entire
history. Police forces themselves have their philosophical origin in
Patrick Colquhoun's 1797 "Treatise on the Police of the Metropolis."
These ideas first took form in London in 1829, but Colquhoun himself
spent his early career as a British agent for cotton manufacturers. His
conception of policing grew from his time spent enforcing slave codes
and working with slave patrols in Virginia. The first hiring of police
through legislative means was in Boston in 1838, but  before then
"police" simply referred to slave patrols. North Carolina's first state
police force was formed in order to stamp out the publishing of "The
Liberator," William Lloyd Garrison's weekly abolitionist newspaper, soon
after he was almost killed in a mob attack in Boston in 1835. The modern
American police force began and grew from these dark roots of American
history.

The existing literature around racialized policing has established that
Black, Hispanic and American Indian/Native Americans are
disproportionately likely to be killed by police in America, especially
at younger ages. One study using data from between 2013 and 2018 found
that Black men are \~2.5 times more and Black women \~1.4 times more
likely to be killed by police in their lifetime than white men and white
women, respectively. This amounts to some 96 out of 100,000 Black men
and boys, between 36 and 81 American Indian/Alaskan Native men and boys,
and 53 out of 100,000 Latino men and boys being killed by police over
their lifetime, as compared to about 39 out of 100,000 white men and
boys being killed. The risk is substantially lower for women across all
racial groups. Between 2.4 and 5.4 Black women and girls, 2.4 American
Indian/Alaskan Native women and girls, and 2 Latino and white women and
girls out of 100,000 are expected to be killed by police throughout
their lifetime. 

In the past decade, the Black Lives Matter Movement rose up in this
context and caused the public to emerge as a strong force to fight for
racial justice and equality in the United States. This movement sparked
a conversation around the appropriate use of force by the police. The
public began pushing for more transparency around the use of force, and
policing general (Schwartz, 2020). That demand has been met, to an
extent, and today we have detailed data on the use of force by many
police departments in the US. In DC, specifically, previous research has
used this research to establish that the majority of citizen-police
interactions involve Black residents and these individuals were much
more likely to be stopped and frisked than white residents (Golash-Boza,
T. Et al, 2022).

This study focuses on data tracking the use of force in 2021 by the
Metropolitan Police Department (MPD) in Washington, DC. MPD is unique in
a number of ways. 51% of its officers were Black and 34% white in 2021
compared to the 46.2% of residents who are white and 45% who are Black.
This makes it one of the few police forces in the US where white
officers are not over represented as compared to the community. This is
commonly thought of as a positive for police forces, and studies have
found that when police diversity matches a neighborhood, crime rates
tend to be lower.

This data includes incidents that occurred between January 1, 2021 and
December 31, 2021. The use of force being defined as any form of
physical effort employed to compel, sway, or convince an individual to
adhere to an officer's directive from the MPD. The reportable forces and
incidents include "Any use of force resulting in injury or a complaint
of injury or pain where the injury or pain is directly associated with a
member's use of force" and "The drawing and pointing of a firearm at, or
in the direction of, another person when no other force was used"
(Metropolitan Department, 2022). The dataset contains 2135 rows and 19
variables, 14 of which are categorical variables and 5 of the others are
quantitative. Variables include "IncidentDate", "IncidentTime",
"IncidentDistrict", CaseStatus", "DateClose", "uof_type", "disposition",
"department_action", "officer_id", "OfficerAssignment" ,
"OfficerGender", "OfficerRace", "year_force", "subject_age",
"subject_race", "subject_gender", "serious", and "CD"(Appendix A). The
dataset is largely complete, with only a few cases of missing data. This
data, extracted from authoritative sources, is a rich resource for
exploring the patterns, implications, and disparities in how force is
applied in the nation's capital.\

In our case, we will be considering the variables  "IncidentDistrict,"
"uof_type," "IncidentDate," "Disposition," "department_action,"
"OfficerGender," "OfficerRace," "subject_age," "subject_race,"
"subject_gender," "serious," and "CD(civil disturbance)." The rest of
the variables (case status, when a case closed, the anonymized officer
IDs, and officer assignment) are either irrelevant or would taint our
data with misleading numbers, such as the anonymized officer IDs.\

**Hypotheses:**

We hypothesize that Black people/African Americans are overrepresented
in use of force cases in Washington, DC. They represented 45% of the DC
population in 2021, so if \> 45% of the subjects of uses of force in our
data are Black/African American, they would be over represented. We also
hypothesize that white police officers are over represented in use of
force cases in Washington, DC. As of 2021, 34.55% of officers were
white. If the % of white officers in use of force cases \> 34.55%, this
would indicate over-representation. This question could serve as a case
study for how accurate representation of a policed community affects the
policing itself.

We are particularly interested in predicting the "disposition," or how
the department classified a use of force, based on the variables
available to us. We hypothesize that uses of force labeled unjustified
may be more common in wards like 7 and 8 that are more diverse and
poorer. The location variable which is available, "IncidentDistrict," is
police district rather than ward, but the police districts are roughly
analogous to the wards:

-   District 1 = Ward 6

-   District 2 = Wards 2 & 3

-   District 3 = Ward 1

-   District 4 = Ward 4

-   District 5 = Ward 5

-   District 6 = Ward 7

-   District 7 = Ward 8

We could do the same type of prediction but for a binary variable like
"serious," which represents the severity of a given use of force.\

**Methods:**

For this study, we will conduct four multiple regressions predicting the
use of force in Washington, DC by the Metropolitan Police Department.
Our predictors will be the police district in which the use of force
took place, the date of the incident, the gender and race of the officer
and the age, race, and gender of the subject. The targets will be the
type of use of force, the disposition, the department action, and the
severity of the force. We won't include the other potential targets as
predictors for one another because severity, for example, will likely be
a very high predictor of the type of use of force, muddling our models.
We will not include variables such as case status, when a case closed,
the anonymized officer IDs, or officer assignments as predictors. We
will remove all uses of force against animals, as we are only interested
in cases involving humans.\

**Data driven hypotheses:**

Our data exploration shows us that Black/African American community
members in DC are subject to a disproportionate amount of force, as
compared to white community members in DC. We can also see that some
wards see more uses of force than others, particularly wards 5, 6, and
7. Our barplot of officer race does seem to indicate that white officers
use force at a higher rate than Black/African American officers, but
this will require further exploration. Our previous hypotheses remain
relevant. We will build multivariate multiple regression models
predicting the type of force, disposition, department action, and
severity of a use of force with the previously discussed variables as
predictors. Given the disproportionate representation of Black/African
Americans in cases of use of force, we expect subject race to be highly
predictive in the model. We also expect the police district, as a proxy
for ward, to be highly predictive, given the disproportionate number of
cases in various districts. We also expect officer race to be
predictive, but to a lesser extent than subject race. We are especially
interested in this variable, given DC's unique positionality as a
majority Black police force.\

**Discussion:**

Our research fills a gap in the existing literature by exploring the use
of force in DC specifically, rather than in the US as a whole or other
metro areas. Other research has explored the implementation of stop and
frisk in DC, but not the use of force broadly. The history of racialized
policing in America is long, and its roots lay close to Washington, DC. 
MPD is uniquely positioned as a majority Black police department in a
demographically diverse city. By exploring how and which of its officers
use force, on whom, and where we may build an understanding of trends in
American policing as a whole. 

```{r}
library(conflicted)
library(dplyr)
library(tidyverse)
library(ggthemes)
library(ggplot2)
library(tibble)

setwd("/Users/kaylhangarcia/Desktop/DATA_412/data412_project/final_project/data-412-mpd-use-of-force")
list.files()
# https://mpdc.dc.gov/node/1635896
df <- read.csv("/Users/kaylhangarcia/Desktop/DATA_412/data412_project/final_project/data-412-mpd-use-of-force/data/UOF_DC_2020maybe.csv")
#df
```

```{r}
mpd_dat <- df
#head(mpd_dat)
#View(mpd_dat)

ggplot(data = mpd_dat, mapping = aes(x = IncidentDistrict)) + 
  geom_bar() +
    facet_wrap(~uof_type)
```

This boxplot provides an overview of use-of-force incidents across
different districts, categorizing them by the specific type of force
used. From the graph, it is evident that certain categories, namely
"hand controls injury," "pointing of a firearm," "tactical takedown
injury," and "Tactical Takedown: No Injury or Complaint of Pain,"
present outstanding counts. Furthermore, districts 6D and 7D stand out
as having the highest incidence rates, with "Tactical Takedown: No
Injury or Complaint of Pain" almost reaching 150 counts. These findings
suggest that these particular districts(6D and 7D) are more prone to
experiencing use-of-force incidents.

```{r}
ggplot(data = mpd_dat, mapping = aes(x = serious)) + 
  geom_bar() +
    facet_wrap(~subject_race)
```

This plot shows the proportion of serious (1) to non serious (0) uses of
force for each subject race. Serious uses of force include firearm
discharges, uses of force resulting in a serious injury, all head
strikes with an impact weapon, uses of force resulting in
unconsciousness or that create a substantial risk of death or
disfigurement, canine bites, neck restraints, or uses of force which
resulted in death.

```{r}
ggplot(data = mpd_dat, mapping = aes(x = subject_race)) + 
  geom_bar()
```

This boxplot takes subject_race as the x-axis and calculates its count
as the y-axis. The data presented in the plot clearly indicates that
Black/African American individuals have experienced a significantly
higher number of use-of-force incidents compared to the other four
racial categories. Conversely, Asian/Pacific Islander individuals appear
to experience the least instances of use of force. This significant
difference between each race can lead to the assumptions that
Black/African American are mostly being targeted by the police.

```{r}
force_by_district_race <- mpd_dat %>%
group_by(OfficerAssignment, subject_race) %>% summarize(Total_Use_of_Force = n()) #total num of cases of use of force

ggplot(subset(force_by_district_race, OfficerAssignment %in% c("1D", "2D", "3D", "4D", "5D", "6D", "7D")), 
       aes(x = OfficerAssignment, y = Total_Use_of_Force, fill = subject_race)) +
  geom_bar(stat = "identity") +
  labs(title = "Use of Force Incidents by District") +
  xlab("Districts") +
  ylab("Total Use of Force Incidents")
```

```{r}
#what dispositions are we most interested in knowing more about?
```

```{r}
#graph for gender and uof type 
mpd_dat <- mpd_dat[complete.cases(mpd_dat), ] #removed nulls

ggplot(mpd_dat, aes(x = subject_gender, fill = uof_type)) +
  geom_bar() +
  labs(title = "Gender and Use of Force Type",
       x = "Gender",
       y = "Count") +
  scale_fill_discrete(name = "UOF Type")

```

This plot shows the number of uses of force and the subject race of
those uses for each police district. Based on the graph, the district of
7D occurs the most incidents of the use of force. For the overall trend,
Black/African American still occurs as the significance in this case.

```{r}
unique(mpd_dat$disposition)
#rename 
#Justified - tactical improvement opp -> jtip 
#Justified - within department policy -> jwdp
#not justified - not within departmental policy -> njnwdp

mpd_dat$disposition <- recode(mpd_dat$disposition,
  "Justified - Tactical Improvement Opportunity" = "JTIP",
  "Tracking Only" = "Tracking",
  "Justified - Within Department Policy" = "JWDP",
  "Not Justified - Not Within Departmental Policy" = "NJNWDP",
  "NA" = "Unknown"
)
```

```{r}
unique(mpd_dat$disposition)
```

This plot shows the count for each gender and uses color to indicate the
different use of force type. The count of male was significant, which
likely suggests that men experience the use of force more than females.
At the same time, the type of "Hand & Straight Strikes", "Other
non-traditional weapons", and "Tactical Takedown: No Injury or Complaint
of Pain" happened the most in the incidents of the use of force.

```{r}
unique(mpd_dat$uof_type)
```

```{r}
unique(mpd_dat$OfficerAssignment)
```

```{r}
districts <- c("1D", "2D", "3D", "4D", "5D", "6D", "7D")
mpd_dat_filtered <- mpd_dat[mpd_dat$OfficerAssignment %in% districts, ]


ggplot(mpd_dat_filtered, aes(x = disposition)) +
  geom_bar() +
  labs(title = "Dispositions for Each District",
       x = "Dispositions",
       y = "Count")
```

```{r}
#df with officer assignment, disposition, officer race
officer_by_district <- mpd_dat %>%
  group_by(OfficerAssignment, disposition, OfficerRace)
```

```{r}
summary_data <- officer_by_district %>%
  filter(OfficerAssignment %in% c("1D")) %>%
  group_by(OfficerAssignment, disposition) %>%
  summarise(count = n())

ggplot(summary_data, aes(x = OfficerAssignment, fill = disposition, y = count)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "District 1: Use of Force Disposition by Officers",
       x = "District",
       y = "Count") +
  facet_wrap(~OfficerAssignment, ncol = 2)

```

```{r}
summary_data <- officer_by_district %>%
  filter(OfficerAssignment %in% c("7D")) %>%
  group_by(OfficerAssignment, disposition) %>%
  summarise(count = n())

ggplot(summary_data, aes(x = OfficerAssignment, fill = disposition, y = count)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "District 7: Use of Force Disposition by Officers",
       x = "District",
       y = "Count") +
  facet_wrap(~OfficerAssignment, ncol = 2)
```

```{r}

districts <- c("1D", "2D", "3D", "4D", "5D", "6D", "7D")
mpd_dat_filtered <- mpd_dat[mpd_dat$OfficerAssignment %in% districts, ]

uof_counts <- mpd_dat_filtered %>%
  group_by(OfficerAssignment) %>%
  summarise(Count = n())

ggplot(uof_counts, aes(x = OfficerAssignment, y = Count)) +
  geom_bar(stat = "identity") +
  labs(title = "Use of Force by District",
       x = "District",
       y = "Count")
```

```{r}
summary_data <- officer_by_district %>%
  filter(OfficerAssignment %in% c("1D", "2D", "3D", "4D", "5D", "6D", "7D")) %>%
  group_by(OfficerAssignment, disposition) %>%
  summarise(count = n())


ggplot(summary_data, aes(x = OfficerAssignment, fill = disposition, y = count)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Use of Force Disposition by Officers District",
       x = "Districts",
       y = "Disposition") +
  facet_wrap(~OfficerAssignment, ncol = 2)
```

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\--

**Citation:**

Schwartz G., (2020). Mapping fatal police violence across U.S.
metropolitan areas: Overall 

rates and racial/ ethnic inequities, 2013-2017. EBSCO.
<https://web.p.ebscohost.com/ehost/pdfviewer/pdfviewer?vid=0&sid=6ba04cd8-6484-4883-8e3e-7e0e405082c8%40redis>

\

Golash-Boza, T. Et al, (2022). Broken windows and order-maintenance
policing in gentrifying Washington, DC. Taylor & Francis Online.
<https://doi.org/10.1080/10439463.2022.2085268>

\

Metropolitan Department. (2022). 2020 and 2021 UoF_Explanatory Notesv2. 

Edwards, F., Lee, H., & Esposito, M. (2019). Risk of being killed by
police use of force in the United States by age, race--ethnicity, and
sex. Proceedings of the National Academy of Sciences, 116(34),
16793--16798. <https://doi.org/10.1073/pnas.1821204116>

Report reveals over 1,000 uses of forces by DC Police in 2021 \|
wusa9.com. (n.d.). Retrieved October 17, 2023, from
<https://www.wusa9.com/article/news/local/dc/annual-use-of-force-report-dc-police-released-more-than-800-complaints-filed/65-9bc43d2b-059d-43bd-a785-3bed9d6ca858>

Lepore, J. (2020, July 13). The New Yorker.
<https://www.newyorker.com/magazine/2020/07/20/the-invention-of-the-police#:~:text=It%20is%20also%20often%20said,the%20history%20of%20the%20police>

Keating, D., & Uhrmacher, K. (2020, June 4). In urban areas, police are
consistently much whiter than the people they serve. The Washington
Post.
<https://www.washingtonpost.com/nation/2020/06/04/urban-areas-police-are-consistently-much-whiter-than-people-they-serve/> 

U.S. Census Bureau QuickFacts: District of Columbia. Accessed October
17, 2023. <https://www.census.gov/quickfacts/fact/table/DC/PST045222>

\

Velez, M. B., Lyons, C. J., & Santoro, W. A. (2015). The political
context of the percent black-neighborhood violence link: A multilevel
analysis. Social Problems, 62(1), 93--119.
<https://doi.org/10.1093/socpro/spu005>
