---
title: "AdamDataExp"
author: "Adam Billen"
date: "10/15/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyverse)
library(ggthemes)
library(ggplot2)
library(tibble)
```
In my mind, the most interesting variables are likely:
-IncidentDistrict
-uof_type
-IncidentDate
-Disposition
-department_action
-OfficerGender
-OfficerRace
-subject_age
-subject_race
-subject_gender
-serious

In my mind, we should remove/ignore:
-IncidentTime
-CaseStatus (I don't think whether the case is close should be an interesting predictor of anything)
-DateClosed (Again, how long ago the case closed seemingly doesn't tell us much)
-Officer_id (unless we want to look at individual officers we can ignore this)
-OfficerAssignment (I could be wrong on this one)
-CD (civil disturbance doesn't seem interesting)
-animal (we should just remove uses of force against animals)


Hypotheses:


```{r}
mpd_dat <- read.csv('UoF_mpd_2021_public.csv')
head(mpd_dat)
View(mpd_dat)

ggplot(data = mpd_dat, mapping = aes(x = IncidentDistrict)) + 
  geom_bar() +
    facet_wrap(~uof_type)

ggplot(data = mpd_dat, mapping = aes(x = serious)) + 
  geom_bar() +
    facet_wrap(~subject_race)

ggplot(data = mpd_dat, mapping = aes(x = subject_race)) + 
  geom_bar()

ggplot(data = mpd_dat, mapping = aes(x = subject_race)) + 
  geom_bar() +
    facet_wrap(~OfficerRace)




```

