---
title: "AdamDataExp"
author: "Adam Billen"
date: "10/15/2023"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

# MPD Use of Force Data Project

```{r}
library(conflicted)
library(dplyr)
library(tidyverse)
library(ggthemes)
library(ggplot2)
library(tibble)
```

In my mind, the most interesting variables are likely: -IncidentDistrict
-uof_type -IncidentDate -Disposition -department_action -OfficerGender
-OfficerRace -subject_age -subject_race -subject_gender -serious -CD
(civil disturbance)

In my mind, we should remove/ignore: -IncidentTime -CaseStatus (I don't
think whether the case is close should be an interesting predictor of
anything) -DateClosed (Again, how long ago the case closed seemingly
doesn't tell us much) -Officer_id (unless we want to look at individual
officers we can ignore this) -OfficerAssignment (I could be wrong on
this one) -animal (we should just remove uses of force against animals)

Hypotheses: We hypothesize that Black people/African Americans are over
represented in use of force cases in Washington, DC. They represented
45% of the DC population in 2021, so if \> 45% of the subjects of uses
of force in our data are Black/African American, they would be over
represented. We also hypothesize that white police officers are over
represented in use of force cases in Washington, DC. As of 2021, 34.55%
of officers were white. If the race of officers in use of force cases \>
34.55%, this would indicate over-representation.

We are particularly interested in predicting the "disposition," or how
the department classified a use of force, based on the variables
available to us. We hypothesize that uses of force labeled unjustified
may be more common in wards 7 and 8 that are more diverse and poorer.
The location variable which is available, "IncidentDistrict," is police
district rather than ward, but we can use the rough lines of the police
district to approximate these areas. We could do the same type of
prediction but for a binary variable like "serious," which represents
the severity of a given use of force.

Methods: Our predictors for these models will be the police district in
which the use of force took place, the date of the incident, the gender
and race of the officer and the age, race, and gender of the subject.
The target wil be the type of use of force, the disposition, the
department action, or the severity of the force. We won't include the
other potential targets as predictors for one another, because severity,
for example, will likely be a very high predictor of the type of use of
force, muddling our models. We will not include variables such as case
status, when a case closed, the anonymized officer IDs, or officer
assignments as predictors. We will remove all uses of force against
animals, as we are only interested in cases involving humans.

```{r}
mpd_dat <- read.csv('UoF_mpd_2021_public.csv')
#head(mpd_dat)
#View(mpd_dat)

ggplot(data = mpd_dat, mapping = aes(x = IncidentDistrict)) + 
  geom_bar() +
    facet_wrap(~uof_type)

```

```{r}
ggplot(data = mpd_dat, mapping = aes(x = serious)) + 
  geom_bar() +
    facet_wrap(~subject_race)

```

```{r}
ggplot(data = mpd_dat, mapping = aes(x = subject_race)) + 
  geom_bar()

```

```{r}
ggplot(data = mpd_dat, mapping = aes(x = subject_race)) + 
  geom_bar() +
    facet_wrap(~OfficerRace)
```
