---
title: "Final Project: MPD Use of Force"
author: "Kaylhan Garcia, Phoebe Lin, Adam Billen"
date: "2023-10-17"
output:
  pdf_document: default
  word_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
library(conflicted)
library(dplyr)
dplyr::filter
conflicts_prefer(dplyr::filter)
```

# MPD Use of Force Final Project Proposal

**Introduction:**

Police have existed in America in some form for nearly its entire
history. Police forces themselves have their philosophical origin in
Patrick Colquhoun's 1797 "Treatise on the Police of the Metropolis."
These ideas first took form in London in 1829, but Colquhoun himself
spent his early career as a British agent for cotton manufacturers. His
conception of policing grew from his time spent enforcing slave codes
and wielding slave patrols in Virginia. The first hiring of police
through legislative means was in Boston in 1838, but  before then
"police" simply referred to slave patrols. North Carolina's first state
police force was formed in order to stamp out the publishing of "The
Liberator," William Lloyd Garrison's weekly abolitionist newspaper, soon
after he was almost killed in a mob attack in Boston in 1835. The modern
American police force began and grew from these dark roots of American
history.

The existing literature has established that Black, Hispanic and
American Indian/Native Americans are disproportionately likely to be
killed by police in America, especially at younger ages. One study using
data from between 2013 and 2018 found that Black men are \~2.5 times
more and Black women \~1.4 times more likely to be killed by police in
their lifetime than white men and white women, respectively. This
amounts to some 96 out of 100,000 Black men and boys, between 36 and 81
American Indian/Alaskan Native men and boys, and 53 out of 100,000
Latino men and boys being killed by police over their lifetime, as
compared to about 39 out of 100,000 white men and boys being killed. The
risk is substantially lower for women across all racial groups. Between
2.4 and 5.4 Black women and girls, 2.4 American Indian/Alaskan Native
women and girls, and 2 Latino and white women and girls out of 100,000
are expected to be killed by police throughout their lifetime. 

In recent years, the Black Lives Matter Movement rose up in this context
and caused the public to emerge as a strong force to fight for racial
justice and equality in the United States. This movement also sparked
the public to think about the appropriation of the use of force from the
police to the general public. Also, the topic of how gender and race can
act upon the use of force. This causes the public to request the
government to provide basic information of the use of force (Schwartz,
2020). Therefore, information about use of force cases becomes more
transparent for the general public to access in order to maintain public
trust. 

Moreover, in the study of Tanya Golash-Boza et al, "Broken windows and
order-maintenance policing in gentrifying Washington, DC", she observed
that the majority of citizen-police interactions were involved with the
Black residents and the Black individuals were even more likely to be
subjected to stop(Golash-Boza, T. Et al, 2022). According to her
findings, this provided us with valuable insights regarding the
demographic group that could be the primary focus of targeting.

For this project, we will be using the dataset about the use of force
that was conducted from the Washington D.C. Metropolitan Police
Department, in other words, the MPD. MPD is unique in a number of ways.
51% of its officers were Black and 34% white in 2021 compared to the
46.2% of residents who are white and 25% who are Black. This makes it
one of the few police forces in the US where white officers are not over
represented as compared to the community. This is commonly thought of as
a positive for police forces, and studies have found that when police
diversity matches a neighborhood, crime rates tend to be lower.\
The dataset focused on the incidents that occurred between January 1,
2021 to December 31, 2021. The Use of force being defined as any form of
physical effort employed to compel, sway, or convince an individual to
adhere to an officer's directive from the MPD. The reportable forces and
incidents include "Any use of force resulting in injury or a complaint
of injury or pain where the injury or pain is directly associated with a
member's use of force" and "The drawing and pointing of a firearm at, or
in the direction of, another person when no other force was used"
(Metropolitan Department, 2022).

The dataset contain 19 variables, 14 of them are categorical variables
and 5 of the others are quantitative. Variables include
**"IncidentDate", "IncidentTime", "IncidentDistrict", CaseStatus",
"DateClose", "uof_type", "disposition", "department_action",
"officer_id", "OfficerAssignment" , "OfficerGender", "OfficerRace",
"year_force", "subject_age", "subject_race", "subject_gender",
"serious",** and "**CD**"(Appendix A).  The categorical variables in
this dataset are mainly ordinal categorical variables. And there are
2135 rows in total. The dataset is mostly completed, only a few boxes
contain "NA". This data, extracted from authoritative sources, is a rich
resource for exploring the patterns, implications, and disparities in
how force is applied in the nation's capital.

This project aims to ...

In our case, we will be considering variables of  *"IncidentDistrict",
"uof_type", "IncidentDate", "Disposition", "department_action",
"OfficerGender", "OfficerRace", "subject_age", "subject_race",
"subject_gender", "serious",* and *"CD(civil disturbance)"* mainly in
our project. Since the rest of the variables wouldn't act significantly
towards our project.

**Hypotheses:**

We hypothesize that Black people/African Americans are overrepresented
in use of force cases in Washington, DC. They represented 45% of the DC
population in 2021, so if \> 45% of the subjects of uses of force in our
data are Black/African American, they would be over represented. We also
hypothesize that white police officers are over represented in use of
force cases in Washington, DC. As of 2021, 34.55% of officers were
white. If the % of white officers in use of force cases \> 34.55%, this
would indicate over-representation. This question could serve as a case
study for how accurate representation of a policed community affects the
policing itself.

We are particularly interested in predicting the "disposition," or how
the department classified a use of force, based on the variables
available to us. We hypothesize that uses of force labeled unjustified
may be more common in wards 7 and 8 that are more diverse and poorer.
The location variable which is available, "IncidentDistrict," is police
district rather than ward, but the police districts are roughly
analogous to the wards:

-   District 1 = Ward 6

-   District 2 = Wards 2 & 3

-   District 3 = Ward 1

-   District 4 = Ward 4

-   District 5 = Ward 5

-   District 6 = Ward 7

-   District 7 = Ward 8

We could do the same type of prediction but for a binary variable like
"serious," which represents the severity of a given use of force.

**Methods:**

Our predictors will be the police district in which the use of force
took place, the date of the incident, the gender and race of the officer
and the age, race, and gender of the subject. The target will be the
type of use of force, the disposition, the department action, or the
severity of the force. We won't include the other potential targets as
predictors for one another, because severity, for example, will likely
be a very high predictor of the type of use of force, muddling our
models. We will not include variables such as case status, when a case
closed, the anonymized officer IDs, or officer assignments as
predictors. We will remove all uses of force against animals, as we are
only interested in cases involving humans.

```{r}
library(conflicted)
library(dplyr)
library(tidyverse)
library(ggthemes)
library(ggplot2)
library(tibble)
# https://mpdc.dc.gov/node/1635896
df <- read.csv("UoF_mpd_2021_public.csv")
#df
```

```{r}
mpd_dat <- read.csv('UoF_mpd_2021_public.csv')
#head(mpd_dat)
#View(mpd_dat)

ggplot(data = mpd_dat, mapping = aes(x = IncidentDistrict)) + 
  geom_bar() +
    facet_wrap(~uof_type)
```

```{r}
ggplot(data = mpd_dat, mapping = aes(x = serious)) + 
  geom_bar() +
    facet_wrap(~subject_race)
```

```{r}
ggplot(data = mpd_dat, mapping = aes(x = subject_race)) + 
  geom_bar()
```

```{r}
ggplot(data = mpd_dat, mapping = aes(x = subject_race)) + 
  geom_bar() +
    facet_wrap(~OfficerRace)
```

```{r}
force_by_district_race <- mpd_dat %>%
  group_by(OfficerAssignment, subject_race) %>%
  summarize(Total_Use_of_Force = n()) #total num of cases of use of force
```

```{r}
ggplot(subset(force_by_district_race, OfficerAssignment %in% c("1D", "2D", "3D", "4D", "5D", "6D", "7D")), 
       aes(x = OfficerAssignment, y = Total_Use_of_Force, fill = subject_race)) +
  geom_bar(stat = "identity") +
  labs(title = "Use of Force Incidents by District") +
  xlab("Districts") +
  ylab("Total Use of Force Incidents")
```

```{r}
#what dispositions are we most interested in knowing more about?
```

```{r}
#graph for gender and uof type 
mpd_dat <- mpd_dat[complete.cases(mpd_dat), ] #removed nulls

ggplot(mpd_dat, aes(x = subject_gender, fill = uof_type)) +
  geom_bar() +
  labs(title = "Gender and Use of Force Type",
       x = "Gender",
       y = "Count") +
  scale_fill_discrete(name = "UOF Type")

```

```{r}
unique(mpd_dat$disposition)
#rename 
#Justified - tactical improvement opp -> jtip 
#Justified - within department policy -> jwdp
#not justified - not within departmental policy -> njnwdp

mpd_dat$disposition <- recode(mpd_dat$disposition,
  "Justified - Tactical Improvement Opportunity" = "JTIP",
  "Tracking Only" = "Tracking",
  "Justified - Within Department Policy" = "JWDP",
  "Not Justified - Not Within Departmental Policy" = "NJNWDP",
  "NA" = "Unknown"
)
```

```{r}
unique(mpd_dat$disposition)
```

```{r}
unique(mpd_dat$uof_type)
```

```{r}
unique(mpd_dat$OfficerAssignment)
```

```{r}
districts <- c("1D", "2D", "3D", "4D", "5D", "6D", "7D")
mpd_dat_filtered <- mpd_dat[mpd_dat$OfficerAssignment %in% districts, ]


ggplot(mpd_dat_filtered, aes(x = disposition)) +
  geom_bar() +
  labs(title = "Dispositions for Each District",
       x = "Dispositions",
       y = "Count")
```

```{r}
#df with officer assignment, disposition, officer race
officer_by_district <- mpd_dat %>%
  group_by(OfficerAssignment, disposition, OfficerRace)
```

```{r}
summary_data <- officer_by_district %>%
  filter(OfficerAssignment %in% c("1D")) %>%
  group_by(OfficerAssignment, disposition) %>%
  summarise(count = n())

ggplot(summary_data, aes(x = OfficerAssignment, fill = disposition, y = count)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "District 1: Use of Force Disposition by Officers",
       x = "District",
       y = "Count") +
  facet_wrap(~OfficerAssignment, ncol = 2)

```

```{r}
summary_data <- officer_by_district %>%
  filter(OfficerAssignment %in% c("7D")) %>%
  group_by(OfficerAssignment, disposition) %>%
  summarise(count = n())

ggplot(summary_data, aes(x = OfficerAssignment, fill = disposition, y = count)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "District 7: Use of Force Disposition by Officers",
       x = "District",
       y = "Count") +
  facet_wrap(~OfficerAssignment, ncol = 2)
```

```{r}

districts <- c("1D", "2D", "3D", "4D", "5D", "6D", "7D")
mpd_dat_filtered <- mpd_dat[mpd_dat$OfficerAssignment %in% districts, ]

uof_counts <- mpd_dat_filtered %>%
  group_by(OfficerAssignment) %>%
  summarise(Count = n())

ggplot(uof_counts, aes(x = OfficerAssignment, y = Count)) +
  geom_bar(stat = "identity") +
  labs(title = "Use of Force by District",
       x = "District",
       y = "Count")
```

```{r}
summary_data <- officer_by_district %>%
  filter(OfficerAssignment %in% c("1D", "2D", "3D", "4D", "5D", "6D", "7D")) %>%
  group_by(OfficerAssignment, disposition) %>%
  summarise(count = n())


ggplot(summary_data, aes(x = OfficerAssignment, fill = disposition, y = count)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Use of Force Disposition by Officers District",
       x = "Districts",
       y = "Disposition") +
  facet_wrap(~OfficerAssignment, ncol = 2)
```

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\--

**Citation:**

Schwartz G., (2020). Mapping fatal police violence across U.S.
metropolitan areas: Overall 

rates and racial/ ethnic inequities, 2013-2017. EBSCO.  

<https://web.p.ebscohost.com/ehost/pdfviewer/pdfviewer?vid=0&sid=6ba04cd8-6484-4883-8e3e-7e0e405082c8%40redis>

\

Golash-Boza, T. Et al, (2022). Broken windows and order-maintenance
policing in gentrifying Washington, DC. Taylor & Francis Online.  

<https://doi.org/10.1080/10439463.2022.2085268>

\

Metropolitan Department. (2022). 2020 and 2021 UoF_Explanatory Notesv2. 

Edwards, F., Lee, H., & Esposito, M. (2019). Risk of being killed by
police use of force in the United States by age, race--ethnicity, and
sex. Proceedings of the National Academy of Sciences, 116(34),
16793--16798. <https://doi.org/10.1073/pnas.1821204116>

Report reveals over 1,000 uses of forces by DC Police in 2021 \|
wusa9.com. (n.d.). Retrieved October 17, 2023, from
<https://www.wusa9.com/article/news/local/dc/annual-use-of-force-report-dc-police-released-more-than-800-complaints-filed/65-9bc43d2b-059d-43bd-a785-3bed9d6ca858>

Lepore, J. (2020, July 13). The New Yorker.
<https://www.newyorker.com/magazine/2020/07/20/the-invention-of-the-police#:~:text=It%20is%20also%20often%20said,the%20history%20of%20the%20police>
