---
title: "mpd project"
author: "Phoebe Lin"
date: "2023-12-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}

library(tidyverse)

```

```{r}
miami <- read_csv("../data/UOF_Miami.csv")
spec(miami)
```

```{r}
miami <- na.omit(miami)
```

Independent: subject race
Dependent: use of force type

```{r}

ggplot(data = miami, mapping = aes(x = Subject_Race)) +
 geom_bar() +
 labs(title = "Histogram of Subject Races", x = "Subject Race", y = "Count")

```

The Histogram of Subject Races displays the distribution of different subject races in the dataset. Based on the histogram, Black and Hispanic are the two highest occurrence in this dataset. There is the trend suggesting that these two racial groups are more likely to be "targeted with use of force"

```{r}
ggplot(data = miami, mapping = aes(x = Type_of_force_used)) +
 geom_bar() +
 labs(title = "Histogram of Types of Force Used", 
      x = "Type of Force Used", 
      y = "Count")

```
This histogram shows the frequency of different types of force used, giving an insight into the most commonly used types of force. In this case, physical force was being used the most in most of the incidents. 

```{r}

ggplot(data = miami, aes(x = Subject_Race, fill = Type_of_force_used)) +
  geom_bar(position = "fill") +
  labs(title = "Stacked Bar Chart of Type of Force Used by Subject Race", 
       x = "Subject Race", 
       y = "Proportion")
```

This stacked bar chart combines the two aspects (subject race and type of force used) to illustrate their relationship. Each bar represents a race, and the different colors within each bar indicate the types of force used for that particular race. This visual helps in understanding if certain types of force are more commonly used with specific races.



