---
title: "Seattle_EDA"
author: "Kaylhan Garcia"
date: "2023-12-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Seattle Exploratory Data Analysis

```{r}
#importing necessary libraries
library(tidyverse)
library(conflicted)
library(dplyr)
library(ggthemes)
library(ggplot2)
library(tibble)
```

```{r}
setwd("/Users/kaylhangarcia/Desktop/DATA_412/data412_project/final_project/data-412-mpd-use-of-force")
list.files()
```

```{r}
library(readr)
seattle <- read_csv("/Users/kaylhangarcia/Desktop/DATA_412/data412_project/final_project/data-412-mpd-use-of-force/data/Use_Of_Force_Seattle.csv")
head(seattle)
```

Finding out all the columns

```{r}
colnames(seattle)
```

```{r}
seattle <- na.omit(seattle)
```

-   He's really emphasising the regressions but also doing exploration on the relationship between our dependent and independent for each variable

-   So it would be most helpful if you guys built ggplots of independent variables vs dependent variables

-   Or if the independent variable isn't continuous then like a histogram vibe

## Predicting Incident Type

Independent: subject race or gender

dependent: incident type

```{r}
unique(seattle$Incident_Type)
```

OIS stands for Office-Involved Shooting

```{r}
ggplot(seattle, aes(x = Subject_Race, fill = Incident_Type)) +
  geom_bar(position = "stack") +
  labs(title = "Distribution of Incident Types by Subject Race",
       x = "Subject Race",
       y = "Count",
       fill = "Incident Type") +
  theme_minimal()
```

```{r}
library(ggplot2)

level3_data <- subset(seattle, Incident_Type %in% c("Level 3 - OIS", "Level 3 - Use of Force"))

ggplot(level3_data, aes(x = Subject_Race, fill = Incident_Type)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Distribution of Level 3 Incidents by Subject Race",
       x = "Subject Race",
       y = "Count",
       fill = "Incident Type") 
```

Interested to see Subject_Race by Precinct and visualize the count and race difference.
