---
title: "Seattle_EDA"
author: "Kaylhan Garcia"
date: "2023-12-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Seattle, WA Exploratory Data Analysis

<https://www.seattle.gov/police/information-and-data/data/use-of-force-data/use-of-force-dataset>

```{r}
#importing necessary libraries
library(tidyverse)
library(conflicted)
library(dplyr)
library(ggthemes)
library(ggplot2)
library(tibble)
```

```{r}
setwd("/Users/kaylhangarcia/Desktop/DATA_412/data412_project/final_project/data-412-mpd-use-of-force")
list.files()
```

```{r}
library(readr)
seattle <- read_csv("/Users/kaylhangarcia/Desktop/DATA_412/data412_project/final_project/data-412-mpd-use-of-force/data/Use_Of_Force_Seattle.csv")
head(seattle)
```

Finding out all the columns

```{r}
colnames(seattle)
```

```{r}
seattle <- na.omit(seattle)
```

-   He's really emphasising the regressions but also doing exploration on the relationship between our dependent and independent for each variable

-   So it would be most helpful if you guys built ggplots of independent variables vs dependent variables

-   Or if the independent variable isn't continuous then like a histogram vibe

## Predicting Incident Type

Independent: subject race or gender

dependent: incident type

```{r}
unique(seattle$Incident_Type)
```

OIS stands for Office-Involved Shooting

```{r}
ggplot(seattle, aes(x = Subject_Race, fill = Incident_Type)) +
  geom_bar(position = "stack") +
  labs(title = "Distribution of Incident Types by Subject Race",
       x = "Subject Race",
       y = "Count",
       fill = "Incident Type") +
  theme_minimal()
```

```{r}
library(ggplot2)

level3_data <- subset(seattle, Incident_Type %in% c("Level 3 - OIS", "Level 3 - Use of Force"))

ggplot(level3_data, aes(x = Subject_Race, fill = Incident_Type)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Distribution of Level 3 Incidents by Subject Race",
       x = "Subject Race",
       y = "Count",
       fill = "Incident Type") 
```

Interested to see Subject_Race by Precinct and visualize the count and race difference.

Sector is smaller than precinct

Beat is more smaller than sector.

```{r}
ggplot(seattle, aes(x = Precinct, fill = Subject_Race)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Distribution of Incidents by Subject Race and Precinct",
       x = "Precinct",
       y = "Count",
       fill = "Subject Race") +
  theme_minimal()
```

```{r}
selected_races <- c("White", "Black or African American", "Hispanic or Latino", "Asian", "Unspecified")
filtered_data <- subset(seattle, Subject_Race %in% selected_races)

ggplot(filtered_data, aes(x = Precinct, fill = Subject_Race)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Distribution of Incidents by Subject Race in South Precinct",
       x = "Precinct",
       y = "Count",
       fill = "Subject Race")
```

Data set includes time range from 2014 (?) to 2023

Population of Seattle: 749,256 (2022)

by Race/Ethnicity:

White alone: 64.9%

Black: 6.8%

American Indian 0.5%

Asian 16.3%

Native Hawaiian 0.2%

Two or more Races 8.8%

Hispanic or Latino 7.2%

White alone, not Hispanic or Latino 62.2%

<https://www.census.gov/quickfacts/fact/table/seattlecitywashington/RHI525222>

-   *West Precinct is where Downtown Seattle is located*

```{r}
ggplot(seattle, aes(x = Sector, fill = Subject_Race)) +
  geom_bar(stat = "count") +
  labs(title = "Distribution of Incidents by Subject Race and Sector",
       x = "Sector",
       y = "Count",
       fill = "Subject Race") +
  theme_minimal()

```

Edward, King, and David Sector are the **highest**

**Lowest**: William, John, and Charles

#### Let's dive deeper into the highest and lowest sectors by incidents

```{r}
highest_sectors <- c("EDWARD", "KING", "DAVID")
lowest_sectors <- c("WILLIAM", "JOHN", "CHARLIE")
```

```{r}
filtered_high_sector <- subset(seattle, Sector %in% highest_sectors)

ggplot(filtered_high_sector, aes(x = Sector, fill = Subject_Race)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Distribution of Incidents by Subject Race in Highest Sectors",
       x = "Sector",
       y = "Count",
       fill = "Subject Race") +
  theme_minimal()
```

\

```{r}
filtered_low_sector <- subset(seattle, Sector %in% lowest_sectors)

ggplot(filtered_low_sector, aes(x = Sector, fill = Subject_Race)) +
  geom_bar(position = "dodge", stat = "count") +
  labs(title = "Distribution of Incidents by Subject Race in Lowest Sectors",
       x = "Sector",
       y = "Count",
       fill = "Subject Race") +
  theme_minimal()
```

```{r}
unique(seattle$Sector)
```

### By Beat 

```{r}
ggplot(seattle, aes(x = Beat, fill = Subject_Race)) +
  geom_bar(stat = "count") +
  labs(title = "Distribution of Incidents by Subject Race and Beat",
       x = "Beat",
       y = "Count",
       fill = "Subject Race") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```

E2 significantly has a higher number of incidents occuring
