---
title: "Use of Force Final Project Report"
author: "Kaylhan Garcia, Adam Billen, and Phoebe Lin"
date: "2023-10-17"
output:
  pdf_document: default
  word_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup}
library(conflicted)
library(tidyverse)
library(nnet)
library(MASS)
library(rpart)
library(coefplot)
conflicts_prefer(dplyr::filter)
```

# Use of Force Final Project Report

**Introduction:**

Police have existed in America in some form for nearly its entire
history. Police forces themselves have their philosophical origin in
Patrick Colquhoun's 1797 "Treatise on the Police of the Metropolis" 
(Colquhoun 1797). These ideas first took form in London in 1829, but 
Colquhoun himself spent his early career as a British agent for cotton manufacturers. His
conception of policing grew from his time spent enforcing slave codes
and working with slave patrols in Virginia. The first hiring of police
through legislative means was in Boston in 1838, but before then
"police" simply referred to slave patrols (Potter 2022). North Carolina's first state
police force was formed in order to stamp out the publishing of "The
Liberator," William Lloyd Garrison's weekly abolitionist newspaper, soon
after he was almost killed in a mob attack in Boston in 1835 (Lepore 2020). The modern
American police force began and grew from these dark roots of American
history.

The existing literature around racialized policing has established that
Black, Hispanic and American Indian/Native Americans are
disproportionately likely to be killed by police in America, especially
at younger ages. One study using data from between 2013 and 2018 found
that Black men are \~2.5 times more and Black women \~1.4 times more
likely to be killed by police in their lifetime than white men and white
women, respectively (Fryer 2019). This amounts to some 96 out of 100,000 Black men
and boys, between 36 and 81 American Indian/Alaskan Native men and boys,
and 53 out of 100,000 Latino men and boys being killed by police over
their lifetime, as compared to about 39 out of 100,000 white men and
boys being killed (Fryer 2019). The risk is substantially lower for women across all
racial groups. Between 2.4 and 5.4 Black women and girls, 2.4 American
Indian/Alaskan Native women and girls, and 2 Latino and white women and
girls out of 100,000 are expected to be killed by police throughout
their lifetime (Fryer 2019). This relationship holds even when controlling for 
uses of police violence against criminal vs non-criminal subjects (Ross et al. 2021). 
This existing set of literature focuses broadly on
fatal violence inflicted by American police, often using national level
data collected by the Washington Post (Washington Post). This data is, however,
scattered and often inconsistent. The Washington Post's database, for example,
relies on local news reports (Nix 2022). Official national level data compiled
by the Bureau of Justice Statistics also fails to provide a full picture. There
is no federal regulation mandating the format or way in which individual police
departments collect data, and so America's 18,000+ police departments each have
control over how, when, and where data is collected. Matthew Hickman, a professor
and former Bureau of Justicts Statistics employee, described the situation as
a "huge mess" wherein police departments can each do "whatever they want" (Katie 
2019). 

In the past decade, however, a new political force, the Black Lives Matter movement,
rose up and sparked a conservation around the appropriate use of force by the
police. The public began pushing for more transparency around the use of force, and
policing generalky (Schwartz, 2020). A number of local governments have, at
least partially, met this demand. These local governments have mandated that
their local police forces provide public data on each case of a police officer
using force against a civilian. The resulting datasets provide a way to analyze
use of force more broadly, rather than soley fatal violence by the police. There
has not, however, been a large wave of research exploring this data. Instead,
local agencies, often Police Accountability Offices, typically produce a yearly
report analyzing the past year's use of force data. While these reports tend to
draw lots of public attention in the news, they are are generally relatively 
surface level. They tend to include general information about the number of uses
of force and any clear disparities in the use of force. They do not, however, 
include more advanced methods such as regression. To expand on this existing
literature, we will emplore more advances methods such as logistic multiple
regression to explore the relationship between demographic factors such as subject
or officer race and the severity of a given use of force. 

**Hypotheses:**

We hypothesize that Black/African American subjects are more likely to have
serious force used against them than other racial groups, particularly
white subjects. While this is our primary topic of interest, we expect that
other demographic groups for both subjects and officers will have varying
degrees of likelihood to have serious force used against them as well. We will also
explore whether particular races of officers are more or less likely to use
serious force against particular races of subjects. We hypthesize that white officers
will be more likely to use serious force against black subjects than black
officers.\

**Methods:**

By including all of
the demographic information provided, we will build a clearer picture of
exactly what demographic groups are more likely to have severe uses of force
inflicted upon them. By subsetting our data to specific officer races, we will
examine which particular combinations of demographic factors result in more 
severe uses of force, providing a fuller picture of how officers and citizens
interact. We have chosen logistic multiple regression as it is the most
applicable and interpretable model for this particular data. All of the data
for each city is categorical, and we have thus created dummy variables for each
key cateogrical variable in each dataset. We have chosen a regular logistic
regression instead of an ordinal logistic regression because each categorical
independent variable is unordered. Including categorical variables such as race
as an ordered set of numbers would make certain races have a larger effect on
the dependent variable, hurting the accuracy and interpretability of the
resulting coefficients. We have created a binary dependent variable (described)
below as our output.

This study involves four unique American cities which vary widely in their
locations and demographic makeup. We will study use of force data in Washington, 
DC (East Coast), Indianapolis (Midwest), Portland (West Coast), and Seattle 
(the Northwest). In order to compare these models we have created a binary
Serious/non-Serious variable for each dataset. This variable is based on a 
pre-existing variable in the dataset for Washington, DC. This variable, "serious,"
is set to 1 when a use of force involves a firearm discharge, strike to the head,
canine bite, neck restraint, results in a serious physical injury, or results in 
the risk of or loss of consciousness, death, or serious disfigurement. The
data for Seattle includes a similar risk tier system from one to three, which
we have simplified to a binary variable where levels two and three are serious
and level one is non-serious. The data for Portland utilizes a four-tiered system,
which we have simplified to make tiers two through four serious uses of force and
tier four non-serious. The data for Indianapolis breaks down uses of force by
individual type, which we have categorized as serious or non-serious to 
match the other cities. Still, these categories should not be read as precisely
comparable. Systems of reporting and publishing data vary between cities, but
this variable still provides a stronger point of reference between cities.\

**Data Exploration**

### Washington, DC

#### 1. Distribution of Incident Counts by Subject Race

Black/ African American have a significantly higher amount of incident counts in 2020. 
Conversely, Asian/Pacific Islander individuals appear to experience the least instances of use of force.

```{r}
df <- read.csv("../data/UOF_DC_2020.csv")
mpd_dat <- df
ggplot(data = mpd_dat, mapping = aes(x = subject_race)) + 
  geom_bar()
```

#### 2. UOF Incident Counts by District and Subject Race

Across all districts in DC, it is prevalent that Black/African Americans experience UOF at a higher rate than their counterparts.

7th District has the highest total UOF incidents out of all the DC districts.

```{r}
force_by_district_race <- mpd_dat %>%
group_by(OfficerAssignment, subject_race) %>% summarize(Total_Use_of_Force = n()) #total num of cases of use of force

ggplot(subset(force_by_district_race, OfficerAssignment %in% c("1D", "2D", "3D", "4D", "5D", "6D", "7D")), 
       aes(x = OfficerAssignment, y = Total_Use_of_Force, fill = subject_race)) +
  geom_bar(stat = "identity") +
  labs(title = "Use of Force Incidents by District") +
  xlab("Districts") +
  ylab("Total Use of Force Incidents")
```

#### 3. Use of Force by District.

1D, 2D, and 3D are shown  to be the safest districts in DC, whereas 7D is deemed to be the most dangerous.

```{r}
districts <- c("1D", "2D", "3D", "4D", "5D", "6D", "7D")
mpd_dat_filtered <- mpd_dat[mpd_dat$OfficerAssignment %in% districts, ]

uof_counts <- mpd_dat_filtered %>%
  group_by(OfficerAssignment) %>%
  summarise(Count = n())

ggplot(uof_counts, aes(x = OfficerAssignment, y = Count)) +
  geom_bar(stat = "identity") +
  labs(title = "Use of Force by District",
       x = "District",
       y = "Count")
```

### Seattle, Washington

```{r}
seattle_2020 <- read_csv("../data/UOF_Seattle.csv")

seattle_2020 <- seattle_2020 %>%
  mutate(datetime = mdy_hms(Occured_date_time)) %>%
  filter(year(datetime) == 2020) %>%
  drop_na(Incident_Type, Subject_Race, Subject_Gender, Precinct, Sector, Beat) %>%
  mutate(serious = ifelse(Incident_Type == "Level 1 - Use of Force", 0, 1))
```

#### 4. Distribution of Incident Types by Subject Race

There is a huge amount of \'Not Specified\' values for this data visualization during the year 2020.

White people have a shared count amount of serious incidents regarding use of force my the police department.

The next following group that stands out is \'Black or African American,\' with level 1 - Use of Force being greater than the other levels.

```{r}
ggplot(seattle_2020, aes(x = Subject_Race, fill = factor(serious))) +
  geom_bar(position = "stack") +
  labs(
    title = "Distribution of Incident Types by Subject Race",
    x = "Subject Race",
    y = "Count",
    fill = "Incident Type"
  ) +
  scale_fill_manual(values = c("0" = "green", "1" = "red"), name = "Serious Incident")
```

#### 5. Comparative Analysis: Incidents in E2 by Subject Race

E2 has the highest amount of incidents across all the other beats in Seattle. There are a total of 52 beats reported in the dataset.

Many of the \'Subject Race\' is left unreported or \'Not Specified\' in the year 2020. The next leading race that are subjected to incidents in the E2 beat are White, then it is Black or African American.

It is notable to mention that many of the serious incidents are reported as \'Not Specified\' in the E2 beat.

```{r}
e2_data <- subset(seattle_2020, Beat == "E2")

ggplot(e2_data, aes(x = Subject_Race, fill = factor(serious))) +
  geom_bar(position = "dodge", stat = "count") +
  labs(
    title = "Distribution of Incidents in E2 Beat",
    x = "Subject Race",
    y = "Count",
    fill = "Serious Incident"
  ) +
  scale_fill_manual(values = c("0" = "green", "1" = "red"), name = "Serious Incident")
```
###Indianapolis, Indiana

In Indianapolis, 52.7% of the population are Whites, 28.5% are Black or African American and 3.82% are Asian. The presence of an 'Unknown' category for subject race suggests that data collection may not be complete or consistently conducted across all incidents. But overall, the "Black" category shows significantly higher number of incidents comparedito other races.

The seriousness of the cases seems to vary across races, but for those races with a significant number of incidents (Black and White categories), non-serious incidents outnumber serious incidents. This could indicate that while the use of force is more commonly reported as non-serious, serious use of force incidents still occur with notable frequency in these populations.

```{r}
ind_dat <- read_csv("../data/ind_dat_2020_final.csv")

ind_dat$serious <- factor(ind_dat$serious)
```


#### Counts of incidents by subject race

```{r}
ggplot(data = ind_dat, mapping = aes(x = RACE)) +
  geom_bar() +
  labs(title = "Counts of incidents by subject race",
       x = "Subject Race",
       y = "Counts")
```
Based on the plot, the Blacks are experiencing the most UOF where over 10,000 incidents happened and followed by the White with counts less than 6,000. 


#### The Distribution of the officer race use of force frequency.

```{r}
ggplot(data = ind_dat, mapping = aes(x = OFF_RACE, fill = RACE)) +
  geom_bar() +
  labs(title = "Distribution of the officer race use of force frequency",
       x = "Office Race",
       y = "Counts")
```
In this case, the White officers are using the most use of force in an incident with highest number of Black being targeted. 

#### The counts of serious incidents relative to subject race. 

```{r}
ggplot(data = ind_dat, mapping = aes(x = RACE, fill = serious)) +
 geom_bar() +
 labs(title = "Histogram of Subject Races and serious incidents", x = "Subject Race", y = "Count")
```
Based on the plot, Black experiencing use of force occurred the most, over 10,000 counts. White experiencing almost less than half of the incidents of the Blacks. 

#### The counts of officer gender use of force frequency in relation to the subject race.

```{r}
ggplot(data = ind_dat, mapping = aes(x = SEX, fill = RACE)) +
 geom_bar() +
 labs(title = "Distribution of officer gender UOF frequency", x = "Officer gender", y = "Count")
```
This plots shows the distribution of the counts of the use of force cases among male and female officers. From the chart, it appears that there are significantly more incidents involving males officer than females officer.

### Portland, Oregon

Portland's population is made up of 635,067 of people where 73.8% are White, 5.6% are Black or African American, 8.5% are Asian. Overall, the Whites are experiencing the highest number of incidents in Portland 2020.  


```{r}
port_dat <- read_csv("../data/pdx_dat_2020_final.csv")
```

#### Histogram of Subject Races and serious incidents

```{r}
port_dat$serious <- factor(port_dat$serious)

ggplot(data = port_dat, mapping = aes(x = Subject__Race, fill = serious)) +
 geom_bar() +
 labs(title = "Histogram of Subject Races and serious incidents", x = "Subject Race", y = "Count")
```

Based on the plot, the race with the highest number of incidents is White, with a notable number of both serious and not serious incidents. This might due to the reason that Portland's population have over 70% Whites with only 5.6% of Blacks or African American. For all races, there are more incidents that are not serious (0) than serious (1), as indicated by the size of the colored sections.

**Data driven hypotheses:**

Our data exploration shows that Black/African American community members across
each city are subject to a disproportionate amount of serious force, as compared
to white community members. We can also see that particular wards and districts
in some cities see a disproprionate amount of use of force generally. Our
previous hypotheses remain relevant. We will use the methods described above
to further explore these relationships.  Given the disproportionate representation 
of Black/African Americans in cases of use of force, we expect subject race 
to be highly predictive in the model. We also expect some locations and 
districts to be highly predictive, as well as officer race.\

**Results**/


### Reading in data
```{r}
mpd_dat_2020 <- read.csv('../data/UOF_DC_2020.csv')
sea_dat_2020 <- read.csv('../data/UOF_Seattle.csv')
pdx_dat_2020 <- read_csv("../data/UOF_Portland.csv")
ind_dat_2020 <- read_csv("../data/UOF_Indianapolis.csv")
```
### Setting up the data for DC
```{r}
age_breaks <- c(18, 25, 35, 45, 55, 65, 100)

mpd_dat_2020 <- mpd_dat_2020 %>%
  mutate(
    subject_age = as.numeric(subject_age),
    AgeBands = cut(subject_age, breaks = age_breaks, labels = c("18_24", "25_34", "35_44", "45_54", "55_65", "66_100"),
                   include.lowest = TRUE, right = FALSE)) %>%
  drop_na(uof_type, IncidentDistrict, OfficerGender, OfficerRace, AgeBands, subject_race, subject_gender) %>%
  mutate(OfficerRace = str_replace_all(OfficerRace, "/", "_")) %>%
  mutate(OfficerRace = str_replace_all(OfficerRace, " ", "_")) %>%
  mutate(subject_race = str_replace_all(subject_race, " ", "_")) %>%
  mutate(subject_race = str_replace_all(subject_race, "/", "_"))

mpd_dat_2020$IncidentDistrict <- factor(mpd_dat_2020$IncidentDistrict)
mpd_dat_2020$IncidentDistrict <- relevel(mpd_dat_2020$IncidentDistrict, ref = "1D")
mpd_dat_2020$OfficerGender <- factor(mpd_dat_2020$OfficerGender)
mpd_dat_2020$OfficerGender <- relevel(mpd_dat_2020$OfficerGender, ref = "Male")
mpd_dat_2020$OfficerRace <- factor(mpd_dat_2020$OfficerRace)
mpd_dat_2020$OfficerRace <- relevel(mpd_dat_2020$OfficerRace, ref = "White_Caucasian")
mpd_dat_2020$AgeBands <- factor(mpd_dat_2020$AgeBands)
mpd_dat_2020$AgeBands <- relevel(mpd_dat_2020$AgeBands, ref = "18_24")
mpd_dat_2020$subject_race <- factor(mpd_dat_2020$subject_race)
mpd_dat_2020$subject_race <- relevel(mpd_dat_2020$subject_race, ref = "White_Caucasian")
mpd_dat_2020$subject_gender <- factor(mpd_dat_2020$subject_gender)
mpd_dat_2020$subject_gender <- relevel(mpd_dat_2020$subject_gender, ref = "Male")

encoded_data_2020 <- model.matrix(~ IncidentDistrict + OfficerGender + OfficerRace + AgeBands + subject_race + subject_gender, data = mpd_dat_2020)

mpd_dat_enc_2020 <- cbind(mpd_dat_2020, encoded_data_2020)
```
### DC overall regression 
The results for our regression classifying serious vs non-serious use of force 
based on location, officer race, subject age, subject race and subject gender
in DC for the year 2020 are shown below. For this regression, districts
are compared to District 1 as the reference, officer and subject race are compared
to white as the reference, subject age is compared to 18-24 as the reference, and
subject gender is compared to male as the reference. The negative coefficient
for subjects aged 55-65, for example, indicates that those in that age bracket
are less likely to have serious force used against them than those in the 18-24
age bracket. These "reference" variables are rows in the dataset where a given
row has all 0s for a given set of variables. This is the result of making our
categorical variables into dummy variables. 

The results of this regression indicate that our hypothesis that Black/African
American subject would be more likely to be subject to a serious use of force
may be incorrect, because the variable for subject race = Black/African American
has a non-positive P value (.3). This does not, however, indicate that Black/African
American residents are less likely to have force used against them than white
subject in general, only that they are not more likely to have serious force
used against them then white subjects. This is important to understand for our
analysis of each of the regressions in this study.

Other variables of interest, however, do show significance. In district 5, for
example, serious force is 1.8% less likely to be used than in district 1 (p = .008).
Subjects aged 25 to 34 are 2.2% less likely to have serious forced used against them
than are subjects aged 18-24 (p = .006). Subjects aged 55 to 65 are 5.1% less likely
to have serious force used against them thatn subjects aged 18-24 (p = .02).

```{r}
model_mpd_2020 <- glm(serious ~ IncidentDistrict2D + IncidentDistrict3D + IncidentDistrict4D + IncidentDistrict5D + IncidentDistrict6D + IncidentDistrictOther + OfficerRaceBlack_African_American + OfficerRaceHispanic + OfficerRaceOther + OfficerGenderFemale + AgeBands25_34 + AgeBands35_44 + AgeBands45_54 + AgeBands55_65 + AgeBands66_100 + subject_raceBlack_African_American + subject_raceHispanic + subject_genderFemale,
               data = mpd_dat_enc_2020)

summary(model_mpd_2020)
coefplot(model_mpd_2020)
```
#### White officers in Washington, DC
This regression utilizes the same variables and setup as the full DC
regression, but includes only white officers.

Our results indicate that white officers in DC are no more likely to use
serious force against one racial group than against another. The subject race 
variables each have insignificant p values. Our results do indicate that white 
officers in DC are 4% less likely to use force in District 5 than in District 1 
(p = .018) force against subjects aged 55 to 65 than against those aged 18-24 (p = .047).

Full results for individual officer race breakdowns for Washington, DC are 
available under "Officer race breakdowns DC" in the appendix.

```{r}
mpd_dat_whiteoff <- mpd_dat_enc_2020 %>%
  filter(OfficerRace == "White_Caucasian")

model_mpd_whiteoff_2020 <- glm(serious ~ IncidentDistrict2D + IncidentDistrict3D + IncidentDistrict4D + IncidentDistrict5D + IncidentDistrict6D + IncidentDistrictOther + OfficerGenderFemale + AgeBands25_34 + AgeBands35_44 + AgeBands45_54 + AgeBands55_65 + subject_raceBlack_African_American + subject_raceHispanic + subject_genderFemale,
               data = mpd_dat_whiteoff)

summary(model_mpd_whiteoff_2020)
coefplot(model_mpd_whiteoff_2020)
```
### Officer + subject race combination counts in DC
Getting the counts
```{r}
mpd_dat_blackoff <- mpd_dat_enc_2020 %>%
  filter(OfficerRace == "Black_African_American")

mpd_dat_hispanicoff <- mpd_dat_enc_2020 %>%
  filter(OfficerRace == "Hispanic")

mpd_dat_whiteoff %>%
  filter(subject_race == "White_Caucasian") %>%
  nrow()

mpd_dat_whiteoff %>%
  filter(subject_race == "Black_African_American") %>%
  nrow()

mpd_dat_whiteoff %>%
  filter(subject_race == "Asian_Pacific_Islander") %>%
  nrow()

mpd_dat_whiteoff %>%
  filter(subject_race == "Hispanic") %>%
  nrow()


mpd_dat_blackoff %>%
  filter(subject_race == "White_Caucasian") %>%
  nrow()

mpd_dat_blackoff %>%
  filter(subject_race == "Black_African_American") %>%
  nrow()

mpd_dat_blackoff %>%
  filter(subject_race == "Asian_Pacific_Islander") %>%
  nrow()

mpd_dat_blackoff %>%
  filter(subject_race == "Hispanic") %>%
  nrow()


mpd_dat_hispanicoff %>%
  filter(subject_race == "White_Caucasian") %>%
  nrow()

mpd_dat_hispanicoff %>%
  filter(subject_race == "Black_African_American") %>%
  nrow()

mpd_dat_hispanicoff %>%
  filter(subject_race == "Asian_Pacific_Islander") %>%
  nrow()

mpd_dat_hispanicoff %>%
  filter(subject_race == "Hispanic") %>%
  nrow()


mpd_dat_hispanicoff %>%
  filter(subject_race == "White_Caucasian") %>%
  nrow()

mpd_dat_hispanicoff %>%
  filter(subject_race == "Black_African_American") %>%
  nrow()

mpd_dat_hispanicoff %>%
  filter(subject_race == "Asian_Pacific_Islander") %>%
  nrow()

mpd_dat_hispanicoff %>%
  filter(subject_race == "Hispanic") %>%
  nrow()
```
The below matrix shows the combination counts of officer + subject race for
each possible combination. We analyzed this relationship by developing
the regressions for specific officer races, as shown above in the regression
for just white officers in DC For the full results of those 
regressions, view "Officer race breakdowns DC" in the appendix.
```{r}
matrix_data_mpd <- matrix(c(
  55, 50, 14,
  800, 757, 181,
  2, 2, 0,
  17, 7, 7
), ncol = 3, byrow = TRUE)

row_names_mpd <- c("White subject", "Black subject", "Hispanic subject", "Asian subject")
col_names_mpd <- c("White officer", "Black officer", "Hispanic officer")

data_df_mpd <- as.data.frame(matrix_data_mpd)
rownames(data_df_mpd) <- row_names_mpd
colnames(data_df_mpd) <- col_names_mpd

print(data_df_mpd)
```

### Setting up the data for Seattle
```{r}
sea_dat_2020 <- sea_dat_2020 %>%
  mutate(datetime = mdy_hms(Occured_date_time)) %>%
  filter(year(datetime) == 2020) %>%
  drop_na(Incident_Type, Subject_Race, Subject_Gender, Precinct, Sector, Beat) %>%
  mutate(serious = ifelse(Incident_Type == "Level 1 - Use of Force", 0, 1)) %>%
  mutate(Subject_Race = str_replace_all(Subject_Race, "/", "_")) %>%
  mutate(Subject_Race = str_replace_all(Subject_Race, " ", "_")) %>%
  mutate(Subject_Gender = str_replace_all(Subject_Gender, " ", "_")) %>%
  mutate(Subject_Gender = str_replace_all(Subject_Gender, "/", "_")) %>%
  mutate(Subject_Gender = str_replace_all(Subject_Gender, "-", "_"))

sea_dat_2020$Subject_Race <- factor(sea_dat_2020$Subject_Race)
sea_dat_2020$Subject_Race <- relevel(sea_dat_2020$Subject_Race, ref = "White")
sea_dat_2020$Subject_Gender <- factor(sea_dat_2020$Subject_Gender)
sea_dat_2020$Subject_Gender <- relevel(sea_dat_2020$Subject_Gender, ref = "Male")
sea_dat_2020$Precinct <- factor(sea_dat_2020$Precinct)
sea_dat_2020$Precinct <- relevel(sea_dat_2020$Precinct, ref = "West")

encoded_seattle_2020 <- model.matrix(~ Subject_Race + Subject_Gender + Precinct, data = sea_dat_2020)

sea_dat_2020_final <- cbind(sea_dat_2020, encoded_seattle_2020)
```
### Overall regression for Seattle
The results for our regression classifying serious vs non-serious use of force
based on subject race, subject, gender, and location for Seattle for the year
2020 are below. For this regression, subject race is compared to white, subject gender
is compared to male, and precinct is compared to the West precinct. These
function the same as described in the DC regression. It is important to note
that key demographic variables such as officer race and gender are not included
in Seattle's reported data, which may affect the results of other variables and
makes this city's results more difficult to compare with the results of other cities.

The results of this regression also indicate that our hypopthesis that Black/
African American subjects would be more likely to be subject to a serious use
of force may be incorrect. Our results indicate that Black/African American 
subjects may be 12.3% less likely to be subject to a serious use of force 
than white subjects (p = 2.45e-07). Our results indicate that Asian subjects
are 16.6% less likely to be subject to a serious use of force than white
subjects (p = .007). American Indian/Alaska Native subjects appear to be 21.6%
less likely to be subject to a serious use of force than white subjects (p = .001).

In terms of other variables, our results indicate that women are 19% less likely
to be subject to a serious use of force than are men (p = 1.35e-15). Finally,
the North, South, OJJ, South, and Southwest precincts each appear to be less likely
to have serious uses of force occur within them than the West precinct. In contrast,
serious uses of force appear to be 17% more likely to occur in the East precinct,
where large-scale protests occured, than in the West precinct (p = 2.77e-14).

```{r}
model_seattle_2020 <- glm(serious ~ Subject_RaceAsian + Subject_RaceBlack_or_African_American + Subject_RaceHispanic_or_Latino + Subject_RaceAmerican_Indian_Alaska_Native + Subject_RaceNat_Hawaiian_Oth_Pac_Islander + Subject_RaceNot_Specified + Subject_RaceTwo_or_More_Races + Subject_GenderFemale + Subject_GenderNon_binary + Subject_GenderOther + Subject_GenderTransgender_Female + Subject_GenderTransgender_Male + Subject_GenderUnknown + Precinct0 + PrecinctEast + PrecinctNorth + PrecinctOOJ + PrecinctSouth + PrecinctSouthwest, data = sea_dat_2020_final)

summary(model_seattle_2020)
coefplot(model_seattle_2020)
```
### Setting up the data for Portland
```{r}
pdx_dat_2020 <- pdx_dat_2020 %>%
  filter(Year == 2020) %>%
  mutate_all(~str_replace_all(., "/", "_")) %>%
  mutate_all(~str_replace_all(., " ", "_")) %>%
  mutate_all(~str_replace_all(., "-", "_")) %>%
  mutate_all(~str_replace_all(., "\\(|\\)", "")) %>%
  rename_all(~str_replace_all(., " ", "_")) %>%
  rename_all(~str_replace_all(., "-", "")) %>%
  rename_all(~str_replace_all(., "/", "_")) %>%
  mutate(Subject_Age = as.numeric(Subject_Age),
    AgeBands = cut(Subject_Age, breaks = age_breaks, labels = c("18_24", "25_34", "35_44", "45_54", "55_65", "66_100"),
                   include.lowest = TRUE, right = FALSE)) %>%
  drop_na(Category_of_Force_Event__Measured_at_Event_Level, Officer_Precinct, Subject__Race, Subject__Sex, Subject__Transient, AgeBands) %>%
  mutate(serious = ifelse(Category_of_Force_Event__Measured_at_Event_Level == "IV", 0, 1))

pdx_dat_2020$Officer_Precinct <- factor(pdx_dat_2020$Officer_Precinct)
pdx_dat_2020$Officer_Precinct <- relevel(pdx_dat_2020$Officer_Precinct, ref = "Central_Precinct")
pdx_dat_2020$Subject__Race <- factor(pdx_dat_2020$Subject__Race)
pdx_dat_2020$Subject__Race <- relevel(pdx_dat_2020$Subject__Race, ref = "White")
pdx_dat_2020$Subject__Sex <- factor(pdx_dat_2020$Subject__Sex)
pdx_dat_2020$Subject__Sex <- relevel(pdx_dat_2020$Subject__Sex, ref = "Male")
pdx_dat_2020$AgeBands <- factor(pdx_dat_2020$AgeBands)
pdx_dat_2020$AgeBands <- relevel(pdx_dat_2020$AgeBands, ref = "18_24")

encoded_pdx_2020 <- model.matrix(~ Officer_Precinct + Subject__Race + Subject__Sex + Subject__Transient + AgeBands, data = pdx_dat_2020)

pdx_dat_2020_final <- cbind(pdx_dat_2020, encoded_pdx_2020)
```
### Full regression for Portland
The results for our regression classifying serious vs non-serious uses of force
based on officer precinct, subject race, subject gender, and subject age
for Portland for the year 2020 are  below. For this regression, officer precinct
is compared to the central precinct, subject race is compared to white, 
subject sex is compared to male, and subject age is compared to the 18-24 bracket.
It is important to note that, as in Seattle, some key demographic variables such as
officer race and gender are not included in the Portland data, which may affect the results of other variables and
makes this city's results more difficult to compare with the results of other cities.

As in Seattle and DC, the results of this regression also indicate that our hypothesis
that Black/African American subjects are more likely to be subject to a serious use
of force may be incorrect. None of the variables for subject race have a 
statistically significant coefficient, indicating that no one racial
group is more likely to have forced used against them than any other in Portland.

Other variables do show clear relationships, however. Our results indicate that 
serious uses of force are more 35% likely to occur in the 
Detectives Precinct (p = .01), 21% more likely
to occur in the East Precinct (p = 1.61e-10), 41.5% more likely to occur in the
K9 precinct (p = .02), and 37% more likely to occur in the TOD precinct 
(p = .00039). Our results also indicate that women are 17% less likely than men
to have serious force used against them (p = 8.36e-07), that subject classified 
as transient are 14% more likely than non-transient subjects to have forced used 
against them (p = 4.50e-07), and that those aged 66 to 100 are 50% less likely
to have force used against them (p = .02).
```{r}
model_portland_2020 <- glm(serious ~ Officer_PrecinctDetectives + Officer_PrecinctEast_Precinct + Officer_PrecinctFamily_Services + Officer_PrecinctK9 + Officer_PrecinctNorth_Precinct + Officer_PrecinctSERT + Officer_PrecinctTOD + Officer_PrecinctTraffic + Officer_PrecinctTraining + Officer_PrecinctTransit_PPB_Only + Subject__RaceBlack + Subject__RaceHispanic + Subject__RaceNative_American + Subject__RaceUnknown + Subject__SexFemale + Subject__TransientUnknown + Subject__TransientYes + AgeBands25_34 + AgeBands35_44 + AgeBands45_54 + AgeBands55_65 + AgeBands66_100, data = pdx_dat_2020_final)

summary(model_portland_2020)
coefplot(model_portland_2020)
```
### Setting up the data for Indianapolis
```{r}
ind_dat_2020 <- read_csv("../data/UOF_Indianapolis.csv")

ind_dat_2020 <- ind_dat_2020 %>%
  mutate_all(~str_replace_all(., "/", "_")) %>%
  mutate_all(~str_replace_all(., " ", "_")) %>%
  mutate_all(~str_replace_all(., ",", "_")) %>%
  mutate_all(~str_replace_all(., "-", "_")) %>%
  mutate_all(~str_replace_all(., "\\(|\\)", "")) %>%
  mutate_all(~str_replace_all(., ":", "_")) %>%
  mutate_all(~str_replace_all(., "__", "_")) %>%
  mutate(DISTRICT = UDTEXT24B) %>%
  mutate(datetime = mdy(OCCURRED_DT)) %>%
  filter(year(datetime) == 2020) %>%
   mutate(OFF_AGE = as.numeric(OFF_AGE),
    OFF_AGE_BANDS = cut(OFF_AGE, breaks = age_breaks, labels = c("18_24", "25_34", "35_44", "45_54", "55_65", "66_100"),
                   include.lowest = TRUE, right = FALSE)) %>%
  mutate(CIT_AGE = as.numeric(CIT_AGE),
    CIT_AGE_BANDS = cut(CIT_AGE, breaks = age_breaks, labels = c("18_24", "25_34", "35_44", "45_54", "55_65", "66_100"),
                   include.lowest = TRUE, right = FALSE)) %>%
  drop_na(RACE, SEX, OFF_RACE, OFF_SEX, OFF_AGE, OFF_AGE_BANDS, CIT_AGE_BANDS, UOF_FORCE_TYPE)

serious_cats <- c("Less_Lethal_Pepperball", "Lethal_Handgun", "Canine_Bite", "Less_Lethal_Bean_Bag", "Physical_Take_Down",
                  "Less_Lethal_CS_OC", "Physical_Kick", "Less_Lethal_CS_Fogger", "Physical_Leg_Sweep",
                  "Physical_Fist_Strike", "Less_Lethal_Taser", "Physical_Knee_Strike", "Physical_Elbow_Strike", 
                  "Physical_Palm_Strike", "Less_Lethal_Clearout_OC", "Lethal_Rifle", "Less_Lethal_Baton", 
                  "Less_Lethal_BPS_Gas", "Less_Lethal_Burning_CS", "Lethal_Shotgun", "Less_Lethal_CS_Grenade")

ind_dat_2020 <- ind_dat_2020 %>% 
  mutate(serious = as.integer(UOF_FORCE_TYPE %in% serious_cats))

ind_dat_2020$RACE <- factor(ind_dat_2020$RACE)
ind_dat_2020$RACE <- relevel(ind_dat_2020$RACE, ref = "White")
ind_dat_2020$SEX <- factor(ind_dat_2020$SEX)
ind_dat_2020$SEX <- relevel(ind_dat_2020$SEX, ref = "Male")
ind_dat_2020$OFF_RACE <- factor(ind_dat_2020$OFF_RACE)
ind_dat_2020$OFF_RACE <- relevel(ind_dat_2020$OFF_RACE, ref = "White")
ind_dat_2020$OFF_SEX <- factor(ind_dat_2020$OFF_SEX)
ind_dat_2020$OFF_SEX <- relevel(ind_dat_2020$OFF_SEX, ref = "Male")
ind_dat_2020$OFF_AGE_BANDS <- factor(ind_dat_2020$OFF_AGE_BANDS)
ind_dat_2020$OFF_AGE_BANDS <- relevel(ind_dat_2020$OFF_AGE_BANDS, ref = "18_24")
ind_dat_2020$CIT_AGE_BANDS <- factor(ind_dat_2020$CIT_AGE_BANDS)
ind_dat_2020$CIT_AGE_BANDS <- relevel(ind_dat_2020$CIT_AGE_BANDS, ref = "18_24")

encoded_ind_2020 <- model.matrix(~ RACE + SEX + OFF_RACE + OFF_SEX + OFF_AGE_BANDS + CIT_AGE_BANDS, data = ind_dat_2020)

ind_dat_2020_final <- cbind(ind_dat_2020, encoded_ind_2020)
```
### Full regression for Indianapolis
The results for our regression classifying serious vs non-serious uses of force
based on subject race, subject sex, subject age, officer race, officer sex, and
officer age for Indiana in 2020 are below. For this regression, race is
compared to white, sex is compared male, and age is compared to 18-24.

Indianapolis is the only city in our analysis where our results match our
initial hypothesis that Black/African American subjects are more likely to be
subject to a serious use of ofrce than white subjects. Our results indicate that 
Black/African American subjects are 6% more likely to be subjected to serious
force than white subjects (p = .006) and that Biracial subjects are 22% more likely
to be subject to serious force than white subjects (p = 3.11e-05).

Other variables also show a clear relationship with serious force. Our results 
indicate that women are 6.5% less likely to be subjected to serious force than 
men (p = .018) and that female officers are 10.5% less likely to use serious force
than male officers (p = .011). Our results also indicate that citizens aged 
45 to 54 are 18% more likely to be subjected to serious force than citizens aged
18-24 (p = 1.14e-07), that officers aged 35 to 44 are 11% more likely to
use serious force than officers aged 18-24, and that officers aged 45 to 54 are
13% more likely to use serious force than officers aged 18-24.
```{r}
model_indianapolis_2020 <- glm(serious ~ RACEBi_racial + RACEBlack + RACEHispanic + RACEUnknown + SEXFemale + OFF_RACEBlack + OFF_RACEHispanic + OFF_RACEOther + OFF_SEXFemale + OFF_AGE_BANDS25_34 + OFF_AGE_BANDS35_44 + OFF_AGE_BANDS45_54 + OFF_AGE_BANDS55_65 + OFF_AGE_BANDS66_100 + CIT_AGE_BANDS25_34 + CIT_AGE_BANDS35_44 + CIT_AGE_BANDS45_54 + CIT_AGE_BANDS55_65, data = ind_dat_2020_final)

summary(model_indianapolis_2020)
coefplot(model_indianapolis_2020)
```
### Officer race subsetted regressions for Indianapolis
#### White officers in Indianapolis
This regression utilizes the same variables and setup as the full Indianapolis
regression, but includes only white officers.

Our results indicate that white officers are 19% more likely to use serious 
force against biracial subjects than white subjects (p = 1.57e-09) and 6.5% more 
likely to use serious force against Black/African American subjects than
white subjects (p = .008). Our results also indicate that white officers are 
12.8% less likely to use force against female subjects than against male 
subjects (p = .0005). Our results also indicate that white officers aged 35 to 
44 are 13% more likely to use serious force than those aged 18-24 (p = .0002), 
as are those aged 34 to 54 (p = .003). Finally, white officers are 17% more
likely to use serious force against citizens aged 45-54 than those aged 18-24 
(p = 8.99e-07) and 13% less likely to use force against those aged 55 to 65 (p= .01).

Full results for individual officer race breakdowns for Indianapolis are available under
"Officer race breakdowns Indianapolis" in the appendix.
```{r}
ind_dat_whiteoff <- ind_dat_2020_final %>%
  filter(OFF_RACE == "White")

model_ind_whiteoff_2020 <- glm(serious ~ RACEBi_racial + RACEBlack + RACEHispanic + RACEUnknown + SEXFemale + OFF_SEXFemale + OFF_AGE_BANDS25_34 + OFF_AGE_BANDS35_44 + OFF_AGE_BANDS45_54 + OFF_AGE_BANDS55_65 + OFF_AGE_BANDS66_100 + CIT_AGE_BANDS25_34 + CIT_AGE_BANDS35_44 + CIT_AGE_BANDS45_54 + CIT_AGE_BANDS55_65, data = ind_dat_whiteoff)

summary(model_ind_whiteoff_2020)
```
### Matrix of officer + subject race combination conuts
Getting the counts
```{r}
ind_dat_blackoff <- ind_dat_2020_final %>%
  filter(OFF_RACE == "Black")

ind_dat_hispanicoff <- ind_dat_2020_final %>%
  filter(OFF_RACE == "Hispanic")


ind_dat_whiteoff %>%
  filter(RACE == "White") %>%
  nrow()

ind_dat_whiteoff %>%
  filter(RACE == "Black") %>%
  nrow()

ind_dat_whiteoff %>%
  filter(RACE == "Hispanic") %>%
  nrow()


ind_dat_blackoff %>%
  filter(RACE == "White") %>%
  nrow()

ind_dat_blackoff %>%
  filter(RACE == "Black") %>%
  nrow()

ind_dat_blackoff %>%
  filter(RACE == "Hispanic") %>%
  nrow()


ind_dat_hispanicoff %>%
  filter(RACE == "White") %>%
  nrow()

ind_dat_hispanicoff %>%
  filter(RACE == "Black") %>%
  nrow()

ind_dat_hispanicoff %>%
  filter(RACE == "Hispanic") %>%
  nrow()
```
The below matrix shows the combination counts of officer + subject race for
each possible combination. We analyzed this relationship by developing
the regressions for specific officer races, as shown above in the regression
for just white officers in Indianapolis. For the fullr results of those 
regressions, view "Officer race breakdowns Indianapolis" in the appendix.
```{r}
matrix_data <- matrix(c(
  601, 1659, 63,
  62, 139, 3,
  36, 100, 3
), ncol = 3, byrow = TRUE)

row_names <- c("White subject", "Black subject", "Hispanic subject")
col_names <- c("White officer", "Black officer", "Hispanic officer")

data_df <- as.data.frame(matrix_data)
rownames(data_df) <- row_names
colnames(data_df) <- col_names

print(data_df)
```
**Discussion:**

Our research fills a gap in the existing literature by using advanved methods
to analyze localized police use of force data, rather than focusing on fatal
use of force based on national data or employing only basic methods, as in 
city level reports. Our results indicate that our hypothesis that Black/African
American subjects are more likely to be subjected to serious force is true for
only one of the cities studied; Indianapolis. Our results also indicate that
for Indianapolis, white officers are particularly likely to use serious force
against Black/African American or Biracial subjects. Beyond subject race, the 
majority of our regressions showed that women and older subjects are less likely
to have force used against them than men and younger subjects. In addition, 
certain precincts and districts in each city showed a higher likelihood for force
to be used. This could be a result of the type of crime or police-subject 
interactions that occur in this part of the city. If, for example, murders are
much more common in one district than another, it would make sense that officers
would be more likely to utilize force in responding to incidents in said district
than in others.

**Limitations**

Our research is limited primarily by the data itself. As is true for national
data, collection and reporting is not standardized at the local level. Different
cities in our study may employ different techniques and standards for when to
report the use of force. It is also possible that the data is influenced by
how likely different groups are to report negative interactions with the police. 
For the data used in this study, if a citizen chooses to report an inappropriate 
use of force by an officer, that complaint will be added to the dataset. If one
racial group or gender is more likely to report negative interactions with the
police, that may skew the data in a way that makes them seem more or less likely
to be subject to various types of force.

**Appendix**
All other regression outputs, including our exploration of individual relationships,
are included below. We did this in order to make the results section smaller
and easier to read.

### Individual variable/dependent relationships for DC.
```{r}
model_mpd_district <- glm(serious ~ IncidentDistrict2D + IncidentDistrict3D + IncidentDistrict4D + IncidentDistrict5D + IncidentDistrict6D + IncidentDistrictOther,
               data = mpd_dat_enc_2020)
summary(model_mpd_district)

model_mpd_offrace <- glm(serious ~ OfficerRaceBlack_African_American + OfficerRaceHispanic + OfficerRaceOther, 
                         data = mpd_dat_enc_2020)
summary(model_mpd_offrace)

model_mpd_offgender <- glm(serious ~ OfficerGenderFemale, data = mpd_dat_enc_2020)
summary(model_mpd_offgender)

model_mpd_age <- glm(serious ~ AgeBands25_34 + AgeBands35_44 + AgeBands45_54 + AgeBands55_65 + AgeBands66_100,
                     data = mpd_dat_enc_2020)
summary(model_mpd_age)

model_mpd_subrace <- glm(serious ~ subject_raceBlack_African_American + subject_raceHispanic,
                     data = mpd_dat_enc_2020)
summary(model_mpd_subrace)

model_mpd_subgender <- glm(serious ~ subject_genderFemale,
                     data = mpd_dat_enc_2020)
summary(model_mpd_subgender)
```
### Officer race breakdowns DC
```{r}
mpd_dat_blackoff <- mpd_dat_enc_2020 %>%
  filter(OfficerRace == "Black_African_American")

model_mpd_blackoff_2020 <- glm(serious ~ IncidentDistrict2D + IncidentDistrict3D + IncidentDistrict4D + IncidentDistrict5D + IncidentDistrict6D + IncidentDistrictOther + OfficerGenderFemale + AgeBands25_34 + AgeBands35_44 + AgeBands45_54 + AgeBands55_65 + AgeBands66_100 + subject_raceBlack_African_American + subject_raceHispanic + subject_genderFemale,
               data = mpd_dat_enc_2020)

summary(model_mpd_blackoff_2020)

mpd_dat_hispanicoff <- mpd_dat_enc_2020 %>%
  filter(OfficerRace == "Hispanic")

model_dat_hispanicoff_2020 <- glm(serious ~ IncidentDistrict2D + IncidentDistrict3D + IncidentDistrict4D + IncidentDistrict5D + IncidentDistrict6D + IncidentDistrictOther + OfficerGenderFemale + AgeBands25_34 + AgeBands35_44 + AgeBands45_54 + AgeBands55_65 + AgeBands66_100 + subject_raceBlack_African_American + subject_raceHispanic + subject_genderFemale,
               data = mpd_dat_hispanicoff)

summary(model_dat_hispanicoff_2020)
```

### Individual variable/dependent relationships for Seattle.
```{r}
model_seattle_subrace <- glm(serious ~ Subject_RaceAsian + Subject_RaceBlack_or_African_American + Subject_RaceHispanic_or_Latino + Subject_RaceAmerican_Indian_Alaska_Native + Subject_RaceNat_Hawaiian_Oth_Pac_Islander + Subject_RaceNot_Specified + Subject_RaceTwo_or_More_Races, data = sea_dat_2020_final)
summary(model_seattle_subrace)

model_seattle_subgender <- glm(serious ~ Subject_GenderFemale + Subject_GenderNon_binary + Subject_GenderOther + Subject_GenderTransgender_Female + Subject_GenderTransgender_Male + Subject_GenderUnknown, data = sea_dat_2020_final)
summary(model_seattle_subgender)

model_seattle_precinct <- glm(serious ~ Precinct0 + PrecinctEast + PrecinctNorth + PrecinctOOJ + PrecinctSouth + PrecinctSouthwest, data = sea_dat_2020_final)
summary(model_seattle_precinct)
```
### Individual variable/dependent relationships for Portland.
```{r}
model_portland_precinct <- glm(serious ~ Officer_PrecinctDetectives + Officer_PrecinctEast_Precinct + Officer_PrecinctFamily_Services + Officer_PrecinctK9 + Officer_PrecinctNorth_Precinct + Officer_PrecinctSERT + Officer_PrecinctTOD + Officer_PrecinctTraffic + Officer_PrecinctTraining + Officer_PrecinctTransit_PPB_Only, data = pdx_dat_2020_final)
summary(model_portland_precinct)

model_portland_race <- glm(serious ~ Subject__RaceBlack + Subject__RaceHispanic + Subject__RaceNative_American + Subject__RaceUnknown, data = pdx_dat_2020_final)
summary(model_portland_race)

model_portland_sex <- glm(serious ~ Subject__SexFemale, data = pdx_dat_2020_final)
summary(model_portland_sex)

model_portland_transient <- glm(serious ~ Subject__TransientUnknown + Subject__TransientYes, data = pdx_dat_2020_final)
summary(model_portland_transient)

model_portland_age <- glm(serious ~ AgeBands25_34 + AgeBands35_44 + AgeBands45_54 + AgeBands55_65 + AgeBands66_100, data = pdx_dat_2020_final)
summary(model_portland_age)
```
## Individual variable/dependent relationships for Indianapolis
```{r}
model_indianapolis_subrace <- glm(serious ~ RACEBi_racial + RACEBlack + RACEHispanic + RACEUnknown, data=ind_dat_2020_final)
summary(model_indianapolis_subrace)

model_indianapolis_subsex <- glm(serious ~ SEXFemale, data=ind_dat_2020_final)
summary(model_indianapolis_subsex)

model_indianapolis_offrace <- glm(serious ~  OFF_RACEBlack + OFF_RACEHispanic + OFF_RACEOther, data=ind_dat_2020_final)
summary(model_indianapolis_offrace)

model_indianapolis_offsex <- glm(serious ~ OFF_SEXFemale, data=ind_dat_2020_final)
summary(model_indianapolis_offsex)

model_indianapolis_offage <- glm(serious ~ OFF_AGE_BANDS25_34 + OFF_AGE_BANDS35_44 + OFF_AGE_BANDS45_54 + OFF_AGE_BANDS55_65 + OFF_AGE_BANDS66_100, data=ind_dat_2020_final)
summary(model_indianapolis_offage)

model_indianapolis_subage <- glm(serious ~ CIT_AGE_BANDS25_34 + CIT_AGE_BANDS35_44 + CIT_AGE_BANDS45_54 + CIT_AGE_BANDS55_65, data=ind_dat_2020_final)
summary(model_indianapolis_subage)
```
## Officer race breakdowns Indianapolis
```{r}
ind_dat_blackoff <- ind_dat_2020_final %>%
  filter(OFF_RACE == "Black")

model_ind_blackoff_2020 <- glm(serious ~ RACEBi_racial + RACEBlack + RACEHispanic + RACEUnknown + SEXFemale + OFF_SEXFemale + OFF_AGE_BANDS25_34 + OFF_AGE_BANDS35_44 + OFF_AGE_BANDS45_54 + OFF_AGE_BANDS55_65 + OFF_AGE_BANDS66_100 + CIT_AGE_BANDS25_34 + CIT_AGE_BANDS35_44 + CIT_AGE_BANDS45_54 + CIT_AGE_BANDS55_65, data = ind_dat_blackoff)

summary(model_ind_blackoff_2020)
coefplot(model_ind_blackoff_2020)


ind_dat_hispanicoff <- ind_dat_2020_final %>%
  filter(OFF_RACE == "Hispanic")

model_ind_hispanicoff_2020 <- glm(serious ~ RACEBi_racial + RACEBlack + RACEHispanic + RACEUnknown + SEXFemale + OFF_SEXFemale + OFF_AGE_BANDS25_34 + OFF_AGE_BANDS35_44 + OFF_AGE_BANDS45_54 + OFF_AGE_BANDS55_65 + OFF_AGE_BANDS66_100 + CIT_AGE_BANDS25_34 + CIT_AGE_BANDS35_44 + CIT_AGE_BANDS45_54 + CIT_AGE_BANDS55_65, data = ind_dat_hispanicoff)

summary(model_ind_hispanicoff_2020)
coefplot(model_ind_hispanicoff_2020)
```

\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-\-

**Citation:**

Schwartz G., (2020). Mapping fatal police violence across U.S. metropolitan areas: Overall.

Fryer, R. G. (2019). Risk of being killed by police use of force in the United States by age, race–ethnicity, and sex. Proceedings of the National Academy of Sciences, 116(34), 16793-16798. doi:10.1073/pnas.1821204116

rates and racial/ ethnic inequities, 2013-2017. EBSCO.
<https://web.p.ebscohost.com/ehost/pdfviewer/pdfviewer?vid=0&sid=6ba04cd8-6484-4883-8e3e-7e0e405082c8%40redis>

Potter, Gary. "The History of Police in America and the First Force." Time Magazine, time.com, 10 June 2022, https://time.com/4779112/police-history-origins/.

Ross, Cody T., et al. "Racial Disparities in Police Use of Deadly Force Against Unarmed Individuals Persist After Appropriately Benchmarking Shooting Data on Violent Crime Rates." Social Psychological and Personality Science, vol. 12, no. 3, 2021, pp. 1948550620916071, doi:10.1177/1948550620916071.

"Police Shootings Database 2015-2023: Search by Race, Age, Department." Washington Post, https://www.washingtonpost.com/graphics/investigations/police-shootings-database/. Accessed 9 Dec. 2023.

Ross, Cody T., Bruce Winterhalder, and Richard McElreath. "Racial Disparities in Police Use of Deadly Force Against Unarmed Individuals Persist After Appropriately Benchmarking Shooting Data on Violent Crime Rates." Social Psychological and Personality Science 12, no. 3 (2021): 1948–5506. doi:10.1177/1948550620916071.

Colquhoun, Patrick. A Treatise on the Police of the Metropolis. 3rd ed., J. Mawman, 1797.

Golash-Boza, T. Et al, (2022). Broken windows and order-maintenance
policing in gentrifying Washington, DC. Taylor & Francis Online.
<https://doi.org/10.1080/10439463.2022.2085268>

Hall, Katie. "Police Use-of-Force Data 'a Huge Mess' Across the U.S." TCA Regional News; Chicago, 8 Sep. 2019.

"The Origins of Modern Day Policing." NAACP, naacp.org, https://naacp.org/find-resources/history-explained/origins-modern-day-policing. Accessed 9 Dec. 2023.

Lepore, Jill. "The Invention of the Police." The New Yorker, 20 July 2020, https://www.newyorker.com/magazine/2020/07/20/the-invention-of-the-police. Accessed 9 Dec. 2023.

Metropolitan Department. (2022). 2020 and 2021 UoF_Explanatory Notesv2. 

Edwards, F., Lee, H., & Esposito, M. (2019). Risk of being killed by
police use of force in the United States by age, race--ethnicity, and
sex. Proceedings of the National Academy of Sciences, 116(34),
16793--16798. <https://doi.org/10.1073/pnas.1821204116>

Report reveals over 1,000 uses of forces by DC Police in 2021
wusa9.com. (n.d.). Retrieved October 17, 2023, from
<https://www.wusa9.com/article/news/local/dc/annual-use-of-force-report-dc-police-released-more-than-800-complaints-filed/65-9bc43d2b-059d-43bd-a785-3bed9d6ca858>

Lepore, J. (2020, July 13). The New Yorker.
<https://www.newyorker.com/magazine/2020/07/20/the-invention-of-the-police#:~:text=It%20is%20also%20often%20said,the%20history%20of%20the%20police>

Keating, D., & Uhrmacher, K. (2020, June 4). In urban areas, police are
consistently much whiter than the people they serve. The Washington
Post.
<https://www.washingtonpost.com/nation/2020/06/04/urban-areas-police-are-consistently-much-whiter-than-people-they-serve/> 

U.S. Census Bureau QuickFacts: District of Columbia. Accessed October
17, 2023. <https://www.census.gov/quickfacts/fact/table/DC/PST045222>

Velez, M. B., Lyons, C. J., & Santoro, W. A. (2015). The political
context of the percent black-neighborhood violence link: A multilevel
analysis. Social Problems, 62(1), 93--119.
<https://doi.org/10.1093/socpro/spu005>